---
title: "BRA_Metabolism"
output: html_document
---

The following script uses StreamMetabolizer to estimate GPP and ER rates at the Brandywine stations.
Installation details. OMIT if necessary
```{r setup, include=FALSE}
remotes::install_github('appling/unitted') #sometimes Rcpp has no binary version yet, if so, don't update from source versions, instead, say no to the source compilation and install the most recent binary version
remotes::install_github("USGS-R/streamMetabolizer") #directly from Git so we get the most recent version

#other necessary packages: 
pack_list<-list ("Rtools","rstan")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}

```

Data input must follow
FOR DO and TEMPREATURE, I used miniDOT data when available and until Jan-7, 2020. Then we switched to DO probe in mayfly.
When data from sensors (miniDOT or Mayfly) are not available or clearly compromised (out of water, malfunction,etc.) I used USGS data to fill gaps
PAR data come from WCD696 station at White Clay Creek
Gage data come from USGS station. 
DOsat concentration is calculated from temperature data prior to analysis
```{r}
library(streamMetabolizer) ; library(unitted) ; library(lubridate) ; library(dplyr)
# file inoput is named BRA_BAM_Met_mmmYYYY
# averaged Do and T data every 15 min, donwloaded PAR from WCD696 (near Stroud), and added gage from Museum USGS station
#time should be in Unix units (output from miniDOT)
#dates are in EST
dat <- read.csv("BRA_BAM_Met_Apr2020.csv",head=T,sep=",") # load csv data
#select only days of baseflow, use the list genrated by FlowSeparation script.
dat$solar.time<-as.POSIXct(dat$solar.time, format="%m/%d/%y %H:%M", tz="EST")
dat$day<- date(dat$solar.time)
dates<-read.csv("BRA_BAM_15min_BaseFlowDates2.csv") ; colnames(dates) <- c("id","day")
dat$day<- as_date(dat$day) ; dates$day <- as_date(dates$day)
dat<- right_join(dat, dates, by="day") ; dat <- dat [,c(-7:-8)]


dat$solar.time<-as.POSIXct(dat$solar.time, format="%m/%d/%y %H:%M", tz="EST")
lubridate::tz(dat$solar.time) # code should be EST
dat$solar.time <- streamMetabolizer::calc_solar_time(dat$solar.time, longitude=-75) #make it mean solar time

unitats<-c(	"", "mgO2 L^-1" ,	"mgO2 L^-1"	, "m"	, "degC" , "umol m^-2 s^-1") #attach units and form a unitted data.frame
dat_unitted<-u(dat, units = unitats)
dat_unitted[c(1,48,96,240,288),] #check that numbers are of right magnitude according to the supposed units 
```

Visualization of raw data
```{r}
library(ggplot2); library(dplyr) ; library(tidyr)
#DO percent saturation and concentration
dat_unitted%>% unitted::v() %>%
  mutate(DO.pctsat = 100 * (DO.obs / DO.sat)) %>%
  select(solar.time, starts_with('DO')) %>%
  gather(type, DO.value, starts_with('DO')) %>%
  mutate(units=ifelse(type == 'DO.pctsat', 'DO\n(% sat)', 'DO\n(mg/L)')) %>%
  ggplot(aes(x=solar.time, y=DO.value, color=type)) + geom_line() + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('variable')

labels <- c(DO.obs='DO concentration\n(mg/L)', DO.sat='DO saturation(%)')
dat_unitted %>% unitted::v() %>%
  select(solar.time, DO.obs, DO.sat) %>%
  gather(type, value, DO.obs, DO.sat) %>%
  mutate(type=ordered(type, levels=c('DO.obs','DO.sat')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line(color='black') + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('black')


#depth,water temperature, and light
labels <- c(depth='depth\n(m)', temp.water='water temp\n(deg C)', light='PAR\n(umol m^-2 s^-1)')
dat_unitted %>% unitted::v() %>%
  select(solar.time, depth, temp.water, light) %>%
  gather(type, value, depth, temp.water, light) %>%
  mutate(
    type=ordered(type, levels=c('depth','temp.water','light')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line() + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('variable')

#water temperature and light
labels <- c( temp.water='water temp\n(deg C)', light='PAR\n(umol m-2 s-1)')
dat_unitted %>% unitted::v() %>%
  select(solar.time, temp.water, light) %>%
  gather(type, value, temp.water, light) %>%
  mutate(
    type=ordered(type, levels=c('temp.water','light')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line(color='black') + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('black')

#DO concentration and saturation, water temperature, and light
labels <- c(DO.obs='DO concentration\n(mg/L)', DO.sat='Do saturation(%)', temp.water='water temp\n(deg C)', light='PAR\n(umol m-2 s-1)')
dat_unitted %>% unitted::v() %>%
  select(DO.obs,DO.sat,solar.time, temp.water, light) %>%
  gather(type, value, DO.obs,DO.sat,temp.water, light) %>%
  mutate(
    type=ordered(type, levels=c('DO.obs','DO.sat','temp.water','light')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line(color='black') + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('black')


```


```{r}
library(Rtools); library(rstan)
#########model 1: Bayesian
bayes_name <- mm_name(type='bayes', pool_K600='none', err_obs_iid=TRUE, err_proc_iid=TRUE)
bayes_name
bayes_specs <- specs(bayes_name)
bayes_specs
#specifications of model (i.e. priors)
bayes_specs <- specs(bayes_name, burnin_steps=500, saved_steps=500, n_cores=4,K600_daily_meanlog = 2.48)
#run model
mm_proba <- metab(bayes_specs, data=dat)
mm <- metab(bayes_specs, data=dat_unitted)
mm
#model parameters
output<-predict_metab(mm_proba)
output
plot_metab_preds(mm_proba)
#check for collinearity between ER and K600
output2<-get_params(mm_proba)
par(mar = c(5,5,1,1))
plot<-plot(output2$K600.daily,output2$ER.daily,xlab =expression("Daily"~K[600]~(d^{-1})),
           pch=16,cex=1.5,col='black',
           ylab=expression("Daily ER" ~ (g ~ O[2]~m^{-2} ~d^{-1})),
           cex.lab=1.2,
           xlim=c(0,60))
output



```



