---
title: "BRA_Metabolism"
output: html_document
---

The following script uses StreamMetabolizer to estimate GPP and ER rates at the Brandywine stations.
Installation details. OMIT if necessary
```{r setup, include=FALSE}
remotes::install_github('appling/unitted') #somteimes Rcpp has no binary version yet, if so, don;t update from source versions, instead, say no to the source compilation and install the most recent binary version
remotes::install_github("USGS-R/streamMetabolizer") #directly from Git so we get the most recent version

#other nececessary packages: 
pack_list<-list ("Rtools","rstan")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}

```

Data input following guidelines:
FOR DO and TEMPREATURE, I used miniDOT data when available and until Jan-7, 2020. Then we switched to DO probe in mayfly.
When data from sensors (miniDOT or Mayfly) are not available or clearly compromised (out of water, malfunction,etc.) I used USGS data to fill gaps
PAR data come from WCD696 station at White Clay Creek
Gage data come from USGS station. 
DOsat concentration is calculated fom temperature data prior to analysis
```{r}
library(streamMetabolizer) ; library(unitted)
# file inoput is named BRA_BAM_Met_mmmYYYY
# averaged Do and T data every 15 min, donwloaded PAR from WCD696 (near Stroud), and added gage from Museum USGS station
#time should be in Unix units (output from miniDOT)
dat <- read.csv("BRA_BAM_Met_Jan2020.csv",head=T,sep=",") # load csv data

dat$solar.time<-as.POSIXct(dat$solar.time, format="%m/%d/%y %H:%M:%S", tz="Etc/GMT+5")
lubridate::tz(dat$solar.time) # code should be EST
dat$solar.time <- streamMetabolizer::calc_solar_time(dat$solar.time, longitude=-75) #make it mean solar time

unitats<-c(	"", "mgO2 L^-1" ,	"mgO2 L^-1"	, "m"	, "degC" , "umol m^-2 s^-1") #attach units and form a unitted data.frame
dat_unitted<-u(dat, units = unitats)
dat_unitted[c(1,48,96,240,288),] #check that numbers are of rigth magnitude according to the suppoesed units 
```

Visualization of raw data
```{r}
library(ggplot2); library(dplyr) ; library(tidyr)
#DO percent saturation and concentration
dat_unitted%>% unitted::v() %>%
  mutate(DO.pctsat = 100 * (DO.obs / DO.sat)) %>%
  select(solar.time, starts_with('DO')) %>%
  gather(type, DO.value, starts_with('DO')) %>%
  mutate(units=ifelse(type == 'DO.pctsat', 'DO\n(% sat)', 'DO\n(mg/L)')) %>%
  ggplot(aes(x=solar.time, y=DO.value, color=type)) + geom_line() + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('variable')

labels <- c(DO.obs='DO concentration\n(mg/L)', DO.sat='DO saturation(%)')
dat_unitted %>% unitted::v() %>%
  select(solar.time, DO.obs, DO.sat) %>%
  gather(type, value, DO.obs, DO.sat) %>%
  mutate(type=ordered(type, levels=c('DO.obs','DO.sat')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line(color='black') + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('black')


#depth,water temperature, and light
labels <- c(depth='depth\n(m)', temp.water='water temp\n(deg C)', light='PAR\n(umol m^-2 s^-1)')
dat_unitted %>% unitted::v() %>%
  select(solar.time, depth, temp.water, light) %>%
  gather(type, value, depth, temp.water, light) %>%
  mutate(
    type=ordered(type, levels=c('depth','temp.water','light')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line() + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('variable')

#water temperature and light
labels <- c( temp.water='water temp\n(deg C)', light='PAR\n(umol m-2 s-1)')
dat_unitted %>% unitted::v() %>%
  select(solar.time, temp.water, light) %>%
  gather(type, value, temp.water, light) %>%
  mutate(
    type=ordered(type, levels=c('temp.water','light')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line(color='black') + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('black')

#DO concetration and saturation, water temperature, and light
labels <- c(DO.obs='DO concentration\n(mg/L)', DO.sat='Do saturation(%)', temp.water='water temp\n(deg C)', light='PAR\n(umol m-2 s-1)')
dat_unitted %>% unitted::v() %>%
  select(DO.obs,DO.sat,solar.time, temp.water, light) %>%
  gather(type, value, DO.obs,DO.sat,temp.water, light) %>%
  mutate(
    type=ordered(type, levels=c('DO.obs','DO.sat','temp.water','light')),
    units=ordered(labels[type], unname(labels))) %>%
  ggplot(aes(x=solar.time, y=value, color=type)) + geom_line(color='black') + 
  facet_grid(units ~ ., scale='free_y') + theme_bw() +
  scale_color_discrete('black')


```


```{r}
remove.packages('rstan')
install.packages('rstan')
library(Rtools); library(rstan)
#########model 1: Bayesian
bayes_name <- mm_name(type='bayes', pool_K600='none', err_obs_iid=TRUE, err_proc_iid=TRUE)
bayes_name
bayes_specs <- specs(bayes_name)
bayes_specs
#specifications of model (i.e. priors)
bayes_specs <- specs(bayes_name, burnin_steps=100, saved_steps=200, n_cores=1,K600_daily_meanlog = 10)
#run model
mm_proba <- metab(bayes_specs, data=dat)
mm <- metab(bayes_specs, data=dat_unitted)
mm
#model parameters
output<-predict_metab(mm_proba)
output
output[c(11:25),]
plot_metab_preds(mm_proba)
#check for collinearity between ER and K600
out<-get_params(mm_proba)
par(mar = c(5,5,1,1))
plot<-plot(output2$K600.daily,output2$ER.daily,xlab =expression("Daily"~K[600]~(d^{-1})),
           pch=16,cex=1.5,col='black',
           ylab=expression("Daily ER" ~ (g ~ O[2]~m^{-2} ~d^{-1})),
           cex.lab=1.2)
output



```

```{r}
######model 2:Maximum likelihood estimation
mle_name<- mm_name(type='mle')
mle_name
mle_specs <- specs(mle_name)
mle_specs
mle_specs <- specs(mle_name, init.GPP.daily=15, init.ER.daily=-12, init.K600.daily=3)
mle_specs <- specs(mle_name, day_start=5, day_end=29, init.K600.daily=100,init.GPP.daily=15)
mle_specs
mm <- metab(mle_specs, data=dat_check2, info=c(site='Miller Creek', source='Kim Bray'))
mm <- metab(mle_specs, data=dat_check2)
#extract specific pieces of info
get_info(mm)
head(get_data(mm))
get_fitting_time(mm) #time for model to run
get_specs(mm) #model specs

#predict and plot metabolism estimates from MLE model
predict_metab(mm)
plot_metab_preds(mm)
#predict and plot DO predictions from MLE model AND observed values
plot_DO_preds(mm)
head(predict_DO(mm))

```

