---
title: "data entry and cleanning"
author: "MPG"
date: "April 5, 2019"
output: html_document 
---
###########################
Package load and retrieval
###########################
```{r,warning=F}
#necessary packages: 
pack_list<-list ("ggplot2","openair","dygraphs","xts","dataRetrieval","dplyr","forecast","glue","dataRetrieval","devtools","DataCombine","lubridate")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}
```

#########################################################################################################
SESTON Data input from .csv and RESEARCH server. 

File is named by sensor's name only.
  The aggregated file and the new one are merged and the time wihtout data due to sensor calibration and   data retrieveal are left as NA

File include date, time, battery, temperature, chl(ug/L), gain voltage (change chl column from scientific format to general)
#########################################################################################################
```{r setup, include=FALSE}
library(dplyr)
setwd("C:/Users/admin/iCloudDrive")
chldata = read.csv("BRA_BAM.csv") ; chldata[,1] <- NULL ; chldata = select (chldata,-c(datetime)) ; tail(chldata,1)
chldata2 = read.csv("BRA_BAM_Jul22-Jul252019.csv") ; head(chldata2,1)
chldata = rbind(chldata,chldata2)
write.csv(chldata, file = "BRA_BAM.csv")
chldata$datetime<- as.vector(as.POSIXct(paste(chldata$date, chldata$time), format="%m/%d/%Y %H:%M:%S", tz="EST"))
```

##################################################################
DISCHARGE data input via URL from USGS using dataRetrieval package
##################################################################
```{r}
library(dataRetrieval)#check station code online (usually an 8-15digit numebr)

###checking parameter code
parameterCdFile <- parameterCdFile
Cds = filter(parameterCdFile, grepl("Discharge", parameterCdFile$parameter_nm,ignore.case=TRUE))

###Input request specifications:
siteNo <- "01481000"
pCode <- "00060"
start.date <- "2019-03-28"
end.date <- "2019-07-25" #----------------------------------------------change end date accordingly

####retrieve data every 15min
        BAM_q <- readNWISuv(siteNumbers = siteNo, 
                            parameterCd = pCode, 
                            startDate = start.date, 
                            endDate = end.date,tz = "EST") 

####retrieve daily average Q
        BAM_q_daily <- readNWISdv(siteNumbers = siteNo, 
                            parameterCd = pCode, 
                            startDate = start.date, 
                            endDate = end.date) 
qdata15 = BAM_q
qdata = BAM_q_daily
```  

######################################################
DAILY TIME SERIES 
1) calculate daily means/max/min from high-frequency data
2) plto daily data ------using dygraphs interactive: 
https://www.r-graph-gallery.com/time-series/-------
######################################################
```{r}
library(xts);library(tidyverse);library(dygraphs);library(forecast);library(timeSeries);library(glue)

chldata$date<-as.Date(chldata$date, format = "%m/%d/%Y") #output is YYY-mm-dd

###Group by date for mean,max,and min. this can be done for Temperature too if needed
      daily_chl_series = chldata %>% 
        group_by(date) %>%
        summarise_at(vars(chl), funs(mean(., na.rm = TRUE), 
                                     (mean(., na.rm = TRUE) + sd(.,na.rm = TRUE)*1.96), 
                                      mean(., na.rm = TRUE) - sd(.,na.rm = TRUE)*1.96))
#interactive plot for daily data
colnames(daily_chl_series) <- c("date","mean","max","min")
  daily_timeseries_chl = xts(x = daily_chl_series[,-1], order.by = daily_chl_series$date)
        dygraph(daily_timeseries_chl,xlab = NULL, ylab = "sestonic chl (ug/L)") %>%
            dySeries(c("min", "mean", "max"))

```

###########################################
DISCHARGE-SESTON RELATIONSHIP
###########################################
```{r}
library(ggplot2); detach(package:dplyr) ; library(dplyr)
qdata2 = qdata[,3:4]; colnames(qdata2)<-c("date","Discharge")
ratingcurve<-merge(daily_chl_series,qdata2,by = c("date"))
ratingcurve_sub = filter(ratingcurve, Discharge <=2000) #To remove abnormal cases or significant outliers

ggplot(ratingcurve_sub, aes(x=log(Discharge), y=log(mean))) +
    geom_point(shape=1,size=3) +    
    geom_smooth(method=lm) +
    xlab("discharge (feet3/s)") +
    ylab("sestonic chl (ug/L)") +
    theme_bw()
```

#########################################################################################################
Moving average of time series
########################################################################################################
```{r}
library(DataCombine)
library(lubridate)

averaged_df<- select(chldata, chl)
a <-as.POSIXct(paste(chldata$date, chldata$time), format="%m/%d/%Y %H:%M:%S", tz="EST")
averaged_df$date <-as.vector(as.POSIXlt(round(as.double(a)/(15*60))*(15*60),origin=(as.POSIXlt('1970-01-01')), tz="EST"))

Q = qdata15[,3:4]; colnames(Q)<-c("date","Discharge")
Q$date <- as.POSIXct(Q$date, format="%m/%d/%Y %H:%M:%S", tz="EST")

PROBA = Q %>% left_join(averaged_df,by = "date")
  
  
  aggregate(x = averaged_df$chl, 
            by = list(unique.dates = averaged_df$dates), 
            FUN = mean)

averaged_df = chldata %>%
  group_by(date) %>%
  summarize(chl = mean(chl, na.rm = TRUE))

aggregate()
head(chldata_mavg, 75)

chl.conc = chldata_mavg[c(-1,-2),8] ; length(chl.conc)
Q = qdata15 ; length(Q)

```

###########################################
DIEL PATTERNS IN SESTON CHL CONCENTRATION
###########################################
```{r}
#subset only baseflow days, threshold can be changed as neededed
                                                                baseflowQ<- 580
chldata_withq = inner_join(chldata,qdata2,by = "date")
chldata_baseflow = filter(chldata_withq, Discharge <=baseflowQ) 
    #count total number of days and prepare plot grid
    dates=as.list(unique(chldata_baseflow$date))
    diels=list() ; Nd<-length(dates)
    par(mfrow = c(round(sqrt(Nd))+1,round(sqrt(Nd))+1))
    par(cex = 0.6)
    par(mar = c(2, 2, 2, 2), oma = c(1, 1, 1, 1))

        #examine diel patterns in chl conc.
          for (i in 1:length(dates)){
            diels[[i]] = filter(chldata_baseflow, date == dates[i])
                 }
                for (i in 1:length(dates)){ 
                  t=diels[[i]]
                  dateN<-t[1,1]
                  plot(t[,5],
                  main=paste0(dateN),
                  ylab="chl(ug/L)",
                  xlab="min since midnight",
                  type="l",
                  col="green")
                                }
```