---
title: "General plots"
author: "MPG"
date: "April 5, 2019"
output: html_document 

README: This script needs to be run when data_entry_script.Rmd has been completed. Order of codechunks matter
---
###########################
Package load and retrieval
###########################
```{r,warning=F}
#necessary packages: 
pack_list<-list ("ggplot2","openair","dygraphs","xts","dataRetrieval","dplyr","forecast","glue","dataRetrieval","devtools","DataCombine","lubridate","zoo")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}
```

###########################################################
Modifiy the following information before running the script
```{r}
qdata = read.csv("BRA_BAM_qdaily.csv")
chldata = read.csv("BRA_BAM.csv")
baseflowQ = 600
stormflowQ = 1000
dataset.15min = read.csv("BRA_BAM_15min.csv")
```

#########################################################
DAILY TIME SERIES 

1) calculate daily means/max/min from high-frequency data
2) plto daily data ------using dygraphs interactive: 
https://www.r-graph-gallery.com/time-series/-------

#########################################################
```{r}
library(xts)
library(tidyverse)
library(dygraphs)
library(forecast)
library(timeSeries)
library(glue)

chldata$date<-as_date(chldata$date) #lubridate function, outut is YYYY-mm-dd

###Group by date for mean,CIup,and CIlow this can be done for Temperature too if needed
      daily_chl_series = chldata %>% group_by(date) %>%
        summarise_at(vars(chl), funs(mean(., na.rm = TRUE), 
                                     (mean(., na.rm = TRUE) + sd(.,na.rm = TRUE)*1.96), 
                                      mean(., na.rm = TRUE) - sd(.,na.rm = TRUE)*1.96))
#interactive plot for daily data
colnames(daily_chl_series) <- c("date","mean","CIup","CIlow")
  daily_timeseries_chl = xts(x = daily_chl_series[,-1], order.by = daily_chl_series$date)
        dygraph(daily_timeseries_chl,xlab = NULL, ylab = "sestonic chl (ug/L)") %>%
            dySeries(c("CIlow", "mean", "CIup"))

```


###########################################
DISCHARGE-SESTON DAILY RELATIONSHIP
###########################################
```{r}
library(ggplot2); library(lubridate)
detach(package:dplyr) ; library(dplyr)

daily_chl_series$date<- as_date(daily_chl_series$date) #lubridate function, outut is YYYY-mm-dd
qdata$date<- as_date(qdata$date) #lubridate function, outut is YYYY-mm-dd
ratingcurve<-inner_join(daily_chl_series,qdata,by = "date")
ratingcurve_sub = filter(ratingcurve, Discharge <=850) #To remove abnormal cases or significant outliers

ggplot(ratingcurve_sub, aes(x=log(Discharge), y=log(mean))) +
    geom_point(shape=1,size=3) +    
    geom_smooth(method=lm) +
    xlab("discharge (feet3/s)") +
    ylab("sestonic chl (ug/L)") +
    theme_bw()
```


###########################################################################################
DIEL PATTERNS IN SESTON CHL CONCENTRATION DURING BASEFLOW
Using 'lubridate' package for time/date data 
The goal is to plot daily patterns for baseflow/stormflow and get a first visual inspection
###########################################################################################
```{r}
#subset only baseflow days from the  'dataset.15min' dataset: value is set on top of the script
chldata_baseflow = filter(dataset.15min, Discharge <=baseflowQ) 
chldata_baseflow = chldata_baseflow %>% 
  tidyr::drop_na(chl.avg)
chldata_baseflow$date<- as_date(chldata_baseflow$date) #lubridate function, output is YYYY-mm-dd. This will allow the loop to subset all the values in a day
    #count total number of days and prepare plot grid
    dates=as.list(unique(chldata_baseflow$date))
    diels=list() ; Nd<-length(dates)
    par(mfrow = c(round(sqrt(Nd)),round(sqrt(Nd))))
    par(cex = 0.6)
    par(mar = c(1, 1, 1, 1), oma = c(1, 1, 1, 1))

        #examine diel patterns in chl conc.
          for (i in 1:length(dates)){
            diels[[i]] = filter(chldata_baseflow, date == dates[i])
                 }
                for (i in 1:length(dates)){ 
                  t=diels[[i]]
                  dateN<-t[1,2]
                  plot(t[,4],
                  main=paste0(dateN),
                  ylab="chl(ug/L)",
                  xlab="min since midnight",
                  type="l",
                  col="green")
                                }
```



###########################################################################################
DIEL PATTERNS IN SESTON CHL CONCENTRATION VS FLOW @BASEFLOW and STORMFLOW

Using 'lubridate' package for time/date data 
The goal is to plot daily patterns for baseflow/stormflow and get a first visual inspection
###########################################################################################
```{r}
#subset only baseflow days from the  'dataset.15min' dataset: value is set on top of the script
chldata_stormflow = filter(dataset.15min, Discharge >=stormflowQ) 
chldata_stormflow = chldata_stormflow %>% 
  tidyr::drop_na(chl.avg)
chldata_stormflow$date<- as_date(chldata_stormflow$date) #lubridate function, output is YYYY-mm-dd. This will allow the loop to subset all the values in a day
    #count total number of days and prepare plot grid
    dates=as.list(unique(chldata_stormflow$date))
    diels=list() ; Nd<-length(dates)
    par(mfrow = c(round(sqrt(Nd)),round(sqrt(Nd))))
    par(cex = 0.6)
    par(mar = c(1, 1, 1, 1), oma = c(1, 1, 1, 1))

        #examine diel patterns in chl conc.
          for (i in 1:length(dates)){
            diels[[i]] = filter(chldata_stormflow, date == dates[i])
                 }
                for (i in 1:length(dates)){ 
                  t=diels[[i]]
                  dateN<-t[1,2]
                  plot(t[,4],
                  main=paste0(dateN),
                  ylab="chl(ug/L)",
                  xlab="min since midnight",
                  type="l",
                  col="green")
                                }

```