---
title: "rawdata input and cleaning"
author: "MPG"
date: "April 5, 2019"
output: html_document 
---
Package loads
```{r,warning=F}
#necessary packages: 
pack_list<-list ("ggplot2","openair","dygraphs","xts","dataRetrieval","dplyr")
for(i in 1:length(pack_list)){
  install.packages(paste("", pack_list[[i]]),repos = "http://cran.us.r-project.org")
}
install.packages("devtools")
install.packages("glue")
install.packages("dataRetrieval", repos=c("http://owi.usgs.gov/R",getOption("repos")))
library(devtools)
install_github("USGS-R/dataRetrieval")
```

data input from .csv and RESEARCH server
file is named by sensor's name only. File include date,time,battery,temperature,chl(ug/L),gain voltage
remember to change chl column from scientific format to general
```{r setup, include=FALSE}
setwd("R:/EcosystemsLab/Ecosystems projects/Brandywine/working_data/sensors/cyclops")
chldata = read.csv("BRA_BAM.csv") ; chldata[,1] <- NULL ; tail(chldata,1)
chldata2 = read.csv("BRA_BAM_May15-Jun262019.csv") ; head(chldata2,1)
chldata = rbind(chldata,chldata2)
write.csv(chldata, file = "BRA_BAM.csv")
```

data input via URL from USGS using dataRetrieval package
```{r}
library(dplyr)
library(dataRetrieval)#check station code online (usually an 8-15digit numebr)

###checking parameter code
parameterCdFile <- parameterCdFile
Cds = filter(parameterCdFile, grepl("Discharge", parameterCdFile$parameter_nm,ignore.case=TRUE))

###Input request specifications:
siteNo <- "01481000"
pCode <- "00060"
start.date <- "2019-03-27"
end.date <- "2019-06-26"

####retrieve data every 15min
BAM_q <- readNWISuv(siteNumbers = siteNo, 
                    parameterCd = pCode, 
                    startDate = start.date, 
                    endDate = end.date) 

####retrieve daily average Q
BAM_q_daily <- readNWISdv(siteNumbers = siteNo, 
                    parameterCd = pCode, 
                    startDate = start.date, 
                    endDate = end.date) 
qdata = BAM_q_daily
```  

Calculate daily means/max/min from high-frequency data
```{r}
library(xts);library(tidyverse)
chldata$date<-as.Date(chldata$date, format = "%m/%d/%Y") #output is YYY-mm-dd
###Group by date for mean,max,and min. this can be done for Temperature too if needed
chl_series = chldata %>% group_by(date) %>%
  summarise_at(vars(chl), funs(mean(., na.rm = TRUE), max(.,na.rm = TRUE), min(.,na.rm = TRUE)))
```

Seston Chlorophyll time series------using dygraphs interactive: https://www.r-graph-gallery.com/time-series/
```{r}
library(dygraphs)
timeseries_chl=xts(x = chl_series[,-1], order.by = chl_series$date)
        dygraph(timeseries_chl,xlab = NULL, ylab = "sestonic chl (ug/L)") %>%
            dySeries(c("min", "mean", "max"))
```

Q-chl rating curve
```{r}
library(ggplot2)
qdata2 = qdata[,3:4]; colnames(qdata2)<-c("date","Discharge")
ratingcurve<-merge(chl_series,qdata2,by = c("date"))
ratingcurve_sub = filter(ratingcurve, mean <= 50 & Discharge <=2000) #To remove abnormal cases or significant outliers

ggplot(ratingcurve_sub, aes(x=Discharge, y=mean)) +
    geom_point(shape=1,size=3) +    
    geom_smooth(method=lm) +
    xlab("discharge (feet3/s)") +
    ylab("sestonic chl (ug/L)") +
    theme_bw()
```

Diel fluctuations in chl & temp
```{r}
chldata_withq = inner_join(chldata,qdata2,by = "date")
chldata_baseflow = filter(chldata_withq, Discharge <=600) #subset only baseflow days
dates=unique(chldata_baseflow$date)
diels=list()

###
for (i in 1:length(dates)){
  diels[[i]] = filter(chldata_baseflow, date == dates[i])
}

###plot baseflow dynamics
for (i in 1:length(dates)){
  t = filter(chldata_baseflow, date == dates[i])
  plot(t[,5])
}




```