---
title: "Seston dynamics"
author: "MPG"
date: "April 5, 2019"
output: html_document 
---
###########################
Package load and retrieval
###########################
```{r,warning=F}
#necessary packages: 
pack_list<-list ("ggplot2","openair","dygraphs","xts","dataRetrieval","dplyr","forecast","glue")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}

install.packages("devtools")
install.packages("dataRetrieval", repos=c("http://owi.usgs.gov/R",getOption("repos")))
library(devtools)
```

#########################################################################################################
SESTON Data input from .csv and RESEARCH server. File is named by sensor's name only. File include date, 
time, battery, temperature, chl(ug/L), gain voltage (change chl column from scientific format to general)
#########################################################################################################
```{r setup, include=FALSE}
library(dplyr)
setwd("R:/EcosystemsLab/Ecosystems projects/Brandywine/working_data/sensors/cyclops")
chldata = read.csv("BRA_BAM.csv") ; chldata %>% select (-c(X)) ; tail(chldata,1)
chldata2 = read.csv("BRA_BAM_Jul22-Jul252019.csv") ; head(chldata2,1)
chldata = rbind(chldata,chldata2)
write.csv(chldata, file = "BRA_BAM.csv")
chldata$datetime<- as.vector(as.POSIXct(paste(chldata$date, chldata$time), format="%m/%d/%Y %H:%M:%S", tz="EST"))
```

##################################################################
DISCHARGE data input via URL from USGS using dataRetrieval package
##################################################################
```{r}
library(dataRetrieval)#check station code online (usually an 8-15digit numebr)

###checking parameter code
parameterCdFile <- parameterCdFile
Cds = filter(parameterCdFile, grepl("Discharge", parameterCdFile$parameter_nm,ignore.case=TRUE))

###Input request specifications:
siteNo <- "01481000"
pCode <- "00060"
start.date <- "2019-03-27"
end.date <- "2019-08-04" #----------------------------------------------change end date accordingly

####retrieve data every 15min
        BAM_q <- readNWISuv(siteNumbers = siteNo, 
                            parameterCd = pCode, 
                            startDate = start.date, 
                            endDate = end.date,tz = "EST") 

####retrieve daily average Q
        BAM_q_daily <- readNWISdv(siteNumbers = siteNo, 
                            parameterCd = pCode, 
                            startDate = start.date, 
                            endDate = end.date) 
qdata15 = BAM_q
qdata = BAM_q_daily
```  

######################################################
DAILY TIME SERIES 
1) calculate daily means/max/min from high-frequency data
2) plto daily data ------using dygraphs interactive: 
https://www.r-graph-gallery.com/time-series/-------
######################################################
```{r}
library(xts);library(tidyverse);library(dygraphs);library(forecast);library(timeSeries);library(glue)

chldata$date<-as.Date(chldata$date, format = "%m/%d/%Y") #output is YYY-mm-dd

###Group by date for mean,max,and min. this can be done for Temperature too if needed
      daily_chl_series = chldata %>% group_by(date) %>%
        summarise_at(vars(chl), funs(mean(., na.rm = TRUE), 
                                     max(.,na.rm = TRUE), 
                                     min(.,na.rm = TRUE)))
#interactive plot for daily data
daily_timeseries_chl = xts(x = daily_chl_series[,-1], order.by = daily_chl_series$date)
        dygraph(daily_timeseries_chl,xlab = NULL, ylab = "sestonic chl (ug/L)") %>%
            dySeries(c("min", "mean", "max"))

```

###########################################
DISCHARGE-SESTON RELATIONSHIP
###########################################
```{r}
library(ggplot2)
qdata2 = qdata[,3:4]; colnames(qdata2)<-c("date","Discharge")
ratingcurve<-merge(daily_chl_series,qdata2,by = c("date"))
ratingcurve_sub = filter(ratingcurve, mean <=50 
                                    & Discharge <=2000) #To remove abnormal cases or significant outliers

ggplot(ratingcurve_sub, aes(x=Discharge, y=mean)) +
    geom_point(shape=1,size=3) +    
    geom_smooth(method=lm) +
    xlab("discharge (feet3/s)") +
    ylab("sestonic chl (ug/L)") +
    theme_bw()
```

###########################################
DIEL PATTERNS IN SESTON CHL CONCENTRATION
###########################################
```{r}
chldata_withq = inner_join(chldata,qdata2,by = "date")
chldata_baseflow = filter(chldata_withq, Discharge <=600) #subset only baseflow days

dates=as.list(unique(chldata_baseflow$date))
diels=list()

###
for (i in 1:length(dates)){
  diels[[i]] = filter(chldata_baseflow, date == dates[i])
}
      for (i in 1:length(dates)){ 
        t=diels[[i]]
        plot(t[,2],t[,5],pch=19,cex=2)
        }


###plot baseflow dynamics
for (i in 1:length(dates)){
  t = filter(chldata_baseflow, date == dates[[i]])
  t$time <- as.POSIXct(strptime(t$time, format="%H:%M:%S"))
        diels[[i]] = ggplot(data = t, aes(x = time, y = chl_avg)) + geom_point(size =2) +
          xlab("Time of Day") +
          ylab("sestonic chl (ug/L)") +
          theme_bw()
}

```