---
title: "data entry and cleanning"
author: "MPG"
date: "April 5, 2019"
output: html_document 

README: Section order is relevant, prompt errors may be caused by failing to run previous code sections
---
###########################
Package load and retrieval
###########################
```{r,warning=F}
#necessary packages: 
pack_list<-list ("ggplot2","openair","dygraphs","xts","dataRetrieval","dplyr","forecast","glue","dataRetrieval","devtools","DataCombine","lubridate","zoo")
for(i in 1:length(pack_list)){
  p <- pack_list[[i]]
  install.packages(p,repos = "https://cran.rstudio.com",dependencies = TRUE)
}
```



#########################################################################################################
SESTON Data input from .csv and RESEARCH server. 

README: File is named by sensor's name only.
  The aggregated file and the new one are merged and the time wihtout data due to sensor calibration and   data retrieveal are left as NA

File include date, time, battery, temperature, chl(ug/L), gain voltage (change chl column from scientific format to general)
#########################################################################################################
```{r setup, include=FALSE}
library(dplyr)

setwd("R:/EcosystemEcologyLab/BRANDYWINEdataDirectory/DataInDevelopment/SensorDeployment/BRA_WILMA_bonnie/04_product")
chldata = read.csv("BRA_WILMA.csv") ; chldata[,1] <- NULL ; chldata = select (chldata,-c(datetime)) ; tail(chldata,1)
setwd("R:/EcosystemEcologyLab/BRANDYWINEdataDirectory/DataInDevelopment/SensorDeployment/BRA_WILMA_bonnie/01_input")
chldata2 = read.csv("BRA_WILMA_Jul22-Jul252019.csv") ; head(chldata2,1)
chldata = rbind(chldata,chldata2)
write.csv(chldata, file = "BRA_WILMA.csv")
chldata$datetime<- as.vector(as.POSIXct(paste(chldata$date, chldata$time), format="%m/%d/%Y %H:%M:%S", tz="EST"))
#chldata contains, among other variables, chlorophyll concentration, temperature, and a Unix datetime column
```



#####################################################################
DISCHARGE data input via URL from USGS station#01481500 in Wilmington 
using dataRetrieval package
#####################################################################
```{r}
library(dataRetrieval)#check station code online (usually an 8-15digit numebr)

###checking parameter code
parameterCdFile <- parameterCdFile
Cds = filter(parameterCdFile, grepl("Discharge", parameterCdFile$parameter_nm,ignore.case=TRUE))

###Input request specifications:
siteNo <- "01481500"
pCode <- "00060"
start.date <- "2019-03-27"
end.date <- "2019-08-04" #----------------------------------------------change end date accordingly

####retrieve data every 15min
        WILMA_q <- readNWISuv(siteNumbers = siteNo, 
                            parameterCd = pCode, 
                            startDate = start.date, 
                            endDate = end.date,tz = "EST") 

####retrieve daily average Q
        WILMA_q_daily <- readNWISdv(siteNumbers = siteNo, 
                            parameterCd = pCode, 
                            startDate = start.date, 
                            endDate = end.date) 
qdata15 = WILMA_q
qdata = WILMA_q_daily
```  



###########################################
DISCHARGE-SESTON RELATIONSHIP
###########################################
```{r}
library(ggplot2)
detach(package:dplyr) ; library(dplyr)

qdata2 = qdata[,3:4]; colnames(qdata2)<-c("date","Discharge")
ratingcurve<-merge(daily_chl_series,qdata2,by = c("date"))
ratingcurve_sub = filter(ratingcurve, Discharge <=2000) #To remove abnormal cases or significant outliers

ggplot(ratingcurve_sub, aes(x=log(Discharge), y=log(mean))) +
    geom_point(shape=1,size=3) +    
    geom_smooth(method=lm) +
    xlab("discharge (feet3/s)") +
    ylab("sestonic chl (ug/L)") +
    theme_bw()
```



#########################################################
DAILY TIME SERIES 

1) calculate daily means/max/min from high-frequency data
2) plto daily data ------using dygraphs interactive: 
https://www.r-graph-gallery.com/time-series/-------

#########################################################
```{r}
library(xts)
library(tidyverse)
library(dygraphs)
library(forecast)
library(timeSeries)
library(glue)

chldata$date<-as.Date(chldata$date, format = "%m/%d/%Y") #output is YYY-mm-dd

###Group by date for mean,max,and min. this can be done for Temperature too if needed
      daily_chl_series = chldata %>% group_by(date) %>%
        summarise_at(vars(chl), funs(mean(., na.rm = TRUE), 
                                     (mean(., na.rm = TRUE) + sd(.,na.rm = TRUE)*1.96), 
                                      mean(., na.rm = TRUE) - sd(.,na.rm = TRUE)*1.96))
#interactive plot for daily data
colnames(daily_chl_series) <- c("date","mean","max","min")
  daily_timeseries_chl = xts(x = daily_chl_series[,-1], order.by = daily_chl_series$date)
        dygraph(daily_timeseries_chl,xlab = NULL, ylab = "sestonic chl (ug/L)") %>%
            dySeries(c("min", "mean", "max"))

```



#########################################################################################################
MOVING AVERAGE SMOOTHING OF ORIGINAL TIME SERIES

1. Generate a merged dataset that contains chl conc. averaged every 15min and paired with discharge data

########################################################################################################
```{r}
library(DataCombine)
library(lubridate)
library(zoo)

######################################15min averaged dataset
averaged_df<- select(chldata, chl)
a <- as.vector(chldata$datetime)
averaged_df$date <-as.vector(as.POSIXlt(round(as.double(a)/(15*60))*(15*60),
                                        origin=(as.POSIXlt('1970-01-01')), 
                                        tz="EST"))

#convert date from 'POSIXlt' class to 'double' to allow aggregation with dplyr
averaged_df$date <- as.double(averaged_df$date) 

#aggregate by date 
dataset.15min = averaged_df %>%
  group_by(date) %>%
  summarize(chl.avg = mean(chl, na.rm = TRUE)) 
#reformat date back to 
dataset.15min$date <- as.POSIXct(dataset.15min$date, format="%m/%d/%Y %H:%M:%S", 
                                           origin=(as.POSIXlt('1970-01-01')),
                                           tz="EST")
#subset so only complete days are selected (3/28/2019 - 8/22/2019)
dataset.15min <- dataset.15min[1:14208,] 

#join with USGS data
Q = qdata15[,3:4]; colnames(Q)<-c("date","Discharge")
#reformat to datetime format 
Q$date <- as.POSIXct(Q$date, format="%m/%d/%Y %H:%M:%S", tz="EST")
#subset so only complete days are selected (3/28/2019 - 8/22/2019)
Q <- Q[5:14204,] ; rownames(Q) <- seq(length=nrow(Q)) #adjust as needed



#IF date column length is not the same we must find which dates are missing
#This takes a while but works. Once identified, use the information to double check the 'right_join' results. The dates identified should have NA in the final dataset
for (i in 1:length(dataset.15min$date)) {
  if(is.element(dataset.15min[i,1],unlist(Q$date))){
    } else {
       print(dataset.15min[i,1]) }
}

dataset.15min <- right_join(Q,dataset.15min,by = "date")
plot(dataset.15min$Discharge,dataset.15min$chl.avg)

```



###########################################################################################
DIEL PATTERNS IN SESTON CHL CONCENTRATION DURING BASEFLOW
Using 'lubridate' package for time/date data 

The goal is to plot daily patterns for baseflow/stormflow and get a first visual inspection
###########################################################################################
```{r}
#######NEED TO merge by 15 min, average by 15min the chl data and temp and merge with USGS data

#subset only baseflow days, threshold can be changed as neededed
                                                                baseflowQ<- 800
chldata_withq = inner_join(chldata,qdata2,by = "date")
chldata_baseflow = filter(chldata_withq, Discharge <=baseflowQ) 
chldata_baseflow = chldata_baseflow[complete.cases(chldata_baseflow), ]
    #count total number of days and prepare plot grid
    dates=as.list(unique(chldata_baseflow$date))
    diels=list() ; Nd<-length(dates)
    par(mfrow = c(round(sqrt(Nd))+1,round(sqrt(Nd))+1))
    par(cex = 0.6)
    par(mar = c(2, 2, 2, 2), oma = c(1, 1, 1, 1))

        #examine diel patterns in chl conc.
          for (i in 1:length(dates)){
            diels[[i]] = filter(chldata_baseflow, date == dates[i])
                 }
                for (i in 1:length(dates)){ 
                  t=diels[[i]]
                  dateN<-t[1,1]
                  plot(t[,5],
                  main=paste0(dateN),
                  ylab="chl(ug/L)",
                  xlab="min since midnight",
                  type="l",
                  col="green")
                                }
```



###########################################################################################
DIEL PATTERNS IN SESTON CHL CONCENTRATION VS FLOW @BASEFLOW and STORMFLOW

Using 'lubridate' package for time/date data 
The goal is to plot daily patterns for baseflow/stormflow and get a first visual inspection

###########################################################################################
```{r}
#flow threshold
                  thresholdQ<- 2000
subset.dataset.15min = filter(dataset.15min, Discharge >=thresholdQ) 
subset.dataset.15min$Date <- as.Date(subset.dataset.15min$date) 
subset.dataset.15min$Time <- format(as.POSIXct(subset.dataset.15min$date) ,format = "%H:%M:%S") 

    #count total number of days and prepare plot grid
    dates=as.list(unique(subset.dataset.15min$Date))
    diels=list() ; Nd<-length(dates)
    par(mfrow = c(round(sqrt(Nd))+1,round(sqrt(Nd))+1))
    par(cex = 0.6)
    par(mar = c(2, 2, 2, 2), oma = c(1, 1, 1, 1))

        #examine diel patterns in chl conc.
          for (i in 1:length(dates)){
            diels[[i]] = filter(subset.dataset.15min, Date == dates[i])
                 }
                for (i in 1:length(dates)){ 
                  t=diels[[i]]
                  dateN<-t[1,1]
                  plot(t[,2], t[,3],
                  main=paste0(dateN),
                  ylab="chl(ug/L)",
                  xlab="discharge",
                  type="p",
                  col="green")
                                }
```